### 2018-07-31 manuscript.rmd
added model adequacy csv, merged Nanomia spp in morphdata

@@

```{r export_table_morph, include=FALSE}
morphdata = cbind(castN, castmeans[,-1], castSE[,-1])
morphdata = morphdata[,c(1:3,33,4,34,5,35,6,36,7,37,8,38,9,39,10,40,11,41,12,42,13,43,14,44,15,45,16,46,17,47,18,48,19,49,20,50,21,51,22,52,23,53,24,54,25,55,26,56,27,57,28,58,29,59,30,60,31,61,32,62)]
names(morphdata) = str_replace_all(names(morphdata),".1","_SE")
write.csv(morphdata, "characterdata.csv")
```

1.05 Plate with all the tentillum diversity.
1.1 Table showing each character’s (illustrating the character per row)  the phylogenetic signal K with the p-value obtained by testing against 1000 simulated permutations, and the number of species included.
```{r phylosignals}
phylosignals = as.data.frame(matrix(ncol=3, nrow=ncol(sharedmean_logs[,-1])))
#phylosignals = as.data.frame(matrix(ncol=3, nrow=ncol(sharedmeans[,-1])))
names(phylosignals) = c("K", "P", "Ntaxa")
for(i in 2:ncol(sharedmean_logs)){
  CH_I=as.numeric(sharedmean_logs[,i])
  #CH_I=as.numeric(sharedmeans[,i])
  names(CH_I) = sharedmean_logs$Species
  #names(CH_I) = sharedmeans$Species
  SE_I = as.numeric(castSE[,i]) %>% log()
  SE_I[which(SE_I==-Inf)] = 0
  names(SE_I) = castSE$Species
  CH_I= CH_I[!is.na(CH_I)]
  SE_I= SE_I[!is.na(SE_I)]
  SE_I = SE_I[which(names(SE_I)%in%names(CH_I))]
  CH_I = CH_I[which(names(CH_I)%in%names(SE_I))]
  treeI = drop.tip(ultram,which(!(ultram$tip.label %in% names(CH_I))))
  class(treeI) = "phylo"
  #print(names(sharedmeans)[i]);print(phylosig(treeI, CH_I, se = SE_I, test=T) %>% .[c("K","P")])
  rownames(phylosignals)[i-1] <- names(sharedmean_logs)[i]
  PSIG <- phylosig(treeI, CH_I, se = SE_I, test=T)
  phylosignals[i-1,1] <- PSIG$K
  phylosignals[i-1,2] <- PSIG$P
  phylosignals[i-1,3] <- length(CH_I)
  phylosignals$K = round(phylosignals$K,3)
}
write.csv(phylosignals, "PhylosignalsWSE_log.csv")
```

@@

```{r fitContinuous}
#Model support
  Q_sharedmean_logs$Heteroneme_to_CB[which(Q_sharedmean_logs$Heteroneme_to_CB > 9)] <- NA
  AICdf = as.data.frame(matrix(ncol=6,nrow=ncol(Q_sharedmean_logs)))
  colnames(AICdf) = c("Variable", "white_noise", "starBM", "BM", "EB", "OU")
  for(c in 1:ncol(Q_sharedmean_logs)){
    startree <- rescale(ultram, "lambda", 0)
    C = Q_sharedmean_logs[,c]
    names(C) = rownames(Q_sharedmean_logs)
    C = C[!is.na(C)]
    Ctree = drop.tip(ultram, which(!(ultram$tip.label %in% names(C))))
    startree = drop.tip(startree, which(!(startree$tip.label %in% names(C))))
    Cse = castSE[,c+1]
    names(Cse) = castSE$Species
    Cse = Cse[which(names(Cse) %in% names(C))]
    model_matrix = matrix("NA", nrow = 5, ncol = 3)
    colnames(model_matrix) = c("aicc","aicc_best","dAICc")
    row.names(model_matrix) = c("white", "starBM", "BM", "EB", "OU")
    for(j in 1:dim(model_matrix)[1]){
      if(j==2){
        temp_model = fitContinuous(startree, C, model="BM", SE = Cse)$opt
      }
      else{
        temp_model = fitContinuous(Ctree, C, model=row.names(model_matrix)[j], SE = Cse)$opt
      }
      model_matrix = apply(model_matrix,2, as.numeric)
      row.names(model_matrix) = c("white", "starBM", "BM", "EB", "OU")
      model_matrix[j, "aicc"] <- temp_model$aicc
    }
    model_matrix[,"aicc_best"] <- min(model_matrix[,"aicc"])
    model_matrix[,"dAICc"] <- model_matrix[, "aicc"] - model_matrix[j, "aicc_best"]
    print(names(Q_sharedmean_logs)[c])
    string_c <- c(names(Q_sharedmean_logs)[c], model_matrix[,3])
    names(string_c) = colnames(AICdf)
    AICdf[c,] <- string_c
  }
  AICdf[,2:6] = apply(AICdf[,2:6], 2, as.numeric) %>% apply(2, round3)
  write.csv(AICdf, "log_model_support.csv")

#Model adequacy
  worthy_models = AICdf[which(AICdf$white_noise != 0 & AICdf$starBM != 0),]
  MAD = as.data.frame(matrix(ncol = 6, nrow = nrow(worthy_models)))
  names(MAD) = c("msig", "cvar", "svar", "sasr", "shgt", "dcfd")
  rownames(MAD) = worthy_models$Variable
  for(m in 1:nrow(worthy_models)){
    C = Q_sharedmean_logs[,which(names(Q_sharedmean_logs) == worthy_models$Variable[m])]
    names(C) = rownames(Q_sharedmean_logs)
    C = C[!is.na(C)]
    Ctree = drop.tip(ultram, which(!(ultram$tip.label %in% names(C))))
    C = C[match(Ctree$tip.label, names(C))]
    class(Ctree)="phylo"
    FC <- fitContinuous(Ctree,C,model=worthy_models$Best_model[m])
    UTC <- make_unit_tree(FC)
    picstat_data <- calculate_pic_stat(UTC)
    sim <- simulate_char_unit(UTC)
    picstat_sim <- calculate_pic_stat(sim)
    compare_pic_stat(picstat_data, picstat_sim) %>% .$p.values -> MAD[m,]
  }
  phy_models = data.frame(worthy_models,MAD)
  phy_models[,7:12] = apply(phy_models[,7:12], 2,round3)
```
1.3 Figure showing major categorical character shifts, gains, and losses on a simplified phylogeny.

>>>>

  #Merge Nanomias
  numbers$Species[which(numbers$Species == "Nanomia cara")] <- "Nanomia sp"
  numbers$Species[which(numbers$Species == "Nanomia bijuga")] <- "Nanomia sp"

@@

```{r export_table_morph, include=FALSE}
morphdata = cbind(castN, castmeans[,-1], castSE[,-1])
morphdata = morphdata[,c(1:3,33,4,34,5,35,6,36,7,37,8,38,9,39,10,40,11,41,12,42,13,43,14,44,15,45,16,46,17,47,18,48,19,49,20,50,21,51,22,52,23,53,24,54,25,55,26,56,27,57,28,58,29,59,30,60,31,61,32,62)]
names(morphdata) = str_replace_all(names(morphdata),".1","_SE")
#write.csv(morphdata, "characterdata.csv")
```

1.05 Plate with all the tentillum diversity.
1.1 Table showing each character’s (illustrating the character per row)  the phylogenetic signal K with the p-value obtained by testing against 1000 simulated permutations, and the number of species included.
```{r phylosignals}
phylosignals = as.data.frame(matrix(ncol=3, nrow=ncol(sharedmean_logs[,-1])))
#phylosignals = as.data.frame(matrix(ncol=3, nrow=ncol(sharedmeans[,-1])))
names(phylosignals) = c("K", "P", "Ntaxa")
for(i in 2:ncol(sharedmean_logs)){
  CH_I=as.numeric(sharedmean_logs[,i])
  #CH_I=as.numeric(sharedmeans[,i])
  names(CH_I) = sharedmean_logs$Species
  #names(CH_I) = sharedmeans$Species
  SE_I = as.numeric(castSE[,i]) %>% log()
  SE_I[which(SE_I==-Inf)] = 0
  names(SE_I) = castSE$Species
  CH_I= CH_I[!is.na(CH_I)]
  SE_I= SE_I[!is.na(SE_I)]
  SE_I = SE_I[which(names(SE_I)%in%names(CH_I))]
  CH_I = CH_I[which(names(CH_I)%in%names(SE_I))]
  treeI = drop.tip(ultram,which(!(ultram$tip.label %in% names(CH_I))))
  class(treeI) = "phylo"
  #print(names(sharedmeans)[i]);print(phylosig(treeI, CH_I, se = SE_I, test=T) %>% .[c("K","P")])
  rownames(phylosignals)[i-1] <- names(sharedmean_logs)[i]
  PSIG <- phylosig(treeI, CH_I, se = SE_I, test=T)
  phylosignals[i-1,1] <- PSIG$K
  phylosignals[i-1,2] <- PSIG$P
  phylosignals[i-1,3] <- length(CH_I)
  phylosignals$K = round(phylosignals$K,3)
}
#write.csv(phylosignals, "PhylosignalsWSE_log.csv")
```

@@

```{r fitContinuous}
#Model support
  Q_sharedmean_logs$Heteroneme_to_CB[which(Q_sharedmean_logs$Heteroneme_to_CB > 9)] <- NA
  AICdf = as.data.frame(matrix(ncol=6,nrow=ncol(Q_sharedmean_logs)))
  colnames(AICdf) = c("Variable", "white_noise", "starBM", "BM", "EB", "OU")
  for(c in 1:ncol(Q_sharedmean_logs)){
    startree <- rescale(ultram, "lambda", 0)
    C = Q_sharedmean_logs[,c]
    names(C) = rownames(Q_sharedmean_logs)
    C = C[!is.na(C)]
    Ctree = drop.tip(ultram, which(!(ultram$tip.label %in% names(C))))
    startree = drop.tip(startree, which(!(startree$tip.label %in% names(C))))
    Cse = castSE[,c+1] %>% log() %>% abs()
    Cse[which(Cse == Inf)] <- 0
    names(Cse) = castSE$Species
    Cse = Cse[which(names(Cse) %in% names(C))]
    model_matrix = matrix("NA", nrow = 5, ncol = 3)
    colnames(model_matrix) = c("aicc","aicc_best","dAICc")
    row.names(model_matrix) = c("white", "starBM", "BM", "EB", "OU")
    for(j in 1:dim(model_matrix)[1]){
      if(j==2){
        temp_model = fitContinuous(startree, C, model="BM", SE = Cse)$opt
      }
      else{
        temp_model = fitContinuous(Ctree, C, model=row.names(model_matrix)[j], SE = Cse)$opt
      }
      model_matrix = apply(model_matrix,2, as.numeric)
      row.names(model_matrix) = c("white", "starBM", "BM", "EB", "OU")
      model_matrix[j, "aicc"] <- temp_model$aicc
    }
    model_matrix[,"aicc_best"] <- min(model_matrix[,"aicc"])
    model_matrix[,"dAICc"] <- model_matrix[, "aicc"] - model_matrix[j, "aicc_best"]
    print(names(Q_sharedmean_logs)[c])
    string_c <- c(names(Q_sharedmean_logs)[c], model_matrix[,3])
    names(string_c) = colnames(AICdf)
    AICdf[c,] <- string_c
  }
  AICdf[,2:6] = apply(AICdf[,2:6], 2, as.numeric) %>% apply(2, round3)
  #write.csv(AICdf, "log_model_support.csv")

#Model adequacy
  worthy_models = AICdf[which(AICdf$white_noise != 0 & AICdf$starBM != 0),]
  MAD = as.data.frame(matrix(ncol = 6, nrow = nrow(worthy_models)))
  names(MAD) = c("msig", "cvar", "svar", "sasr", "shgt", "dcfd")
  rownames(MAD) = worthy_models$Variable
  for(m in 1:nrow(worthy_models)){
    C = Q_sharedmean_logs[,which(names(Q_sharedmean_logs) == worthy_models$Variable[m])]
    names(C) = rownames(Q_sharedmean_logs)
    C = C[!is.na(C)]
    Ctree = drop.tip(ultram, which(!(ultram$tip.label %in% names(C))))
    C = C[match(Ctree$tip.label, names(C))]
    class(Ctree)="phylo"
    FC <- fitContinuous(Ctree,C,model=worthy_models$Best_model[m])
    UTC <- make_unit_tree(FC)
    picstat_data <- calculate_pic_stat(UTC)
    sim <- simulate_char_unit(UTC)
    picstat_sim <- calculate_pic_stat(sim)
    compare_pic_stat(picstat_data, picstat_sim) %>% .$p.values -> MAD[m,]
  }
  phy_models = data.frame(worthy_models,MAD)
  phy_models[,7:12] = apply(phy_models[,7:12], 2,round3)
  #write.csv(phy_models, "model_adequacy.csv")
```
1.3 Figure showing major categorical character shifts, gains, and losses on a simplified phylogeny.
